---
title: 'Find canonical equation'
author: 'fnaufel'
email: 'https://fnaufel.github.io/'
date: '   (v. `r format(Sys.Date(), "%d/%m/%Y")`)'
lang: 'en'

output: 
  # To install these output formats, run
  #   install.packages("devtools")
  #   devtools::install_github("fnaufel/fnaufelRmd")
  fnaufelRmd::html_report:
    []
---

```{r setup, include=FALSE}
# The next command configures MANY things and loads quite a few packages.
# 
# If you want to see what's being done, execute 
# 
#   cat(
#     system.file(
#       "rmarkdown/resources/R/_common_report.R", 
#       package = "fnaufelRmd"
#     )
#   )
# 
# to find out the location of the file. Then open the file.
# 
# If you want to change the configuration, copy the file, edit it, and
# source it instead of the package file. 
# 
# Or simply write your commands here in this code chunk.

source(
  system.file(
    "rmarkdown/resources/R/_common_report.R",
    package = "fnaufelRmd"
  )
)

library(reticulate)
library(sympyglue)

sympy = import('sympy')
py_run_string('from sympy import *')
py_run_string('init_printing(use_latex=True)')
py_run_string('from pprint import pp')
py_run_string('x, y = symbols(\'x y\')')
```


# Complete the square using match

```{python}
def complsq(expr, var):
  
  # expr is of the form ax^2 + bx + c, and var == x
  
  error_msg = '''
  Expression must be of the form ax^2 + bx + c.
  The input was ''' + latex(expr)
  
  # var may be any symbol, but let's call it x here
  x = var
  
  # Used to match the coefficients
  a, b, c = [ \
    Wild(w, exclude=[x, y]) for w in 'abc' \
  ]
  
  # Match
  m = expr.match(a*x**2 + b*x + c)
  
  # No match: bad expression
  if m is None:
    raise ValueError(error_msg)
  
  # Get coefficients
  A, B, C = [w.xreplace(m) for w in [a, b, c]]
  
  # Expression must be of degree 2
  if A == 0:
    raise ValueError(error_msg)
  
  # Will return A (x + B/2A)^2 + (4AC - B^2)/4A
  second = sympify(B) / (2*A)
  third = sympify(4*A*C - B**2) / (4*A)
  
  return A * (var + second)**2 + third
```


## Test complete square

```{python}
A, B, C = symbols('A B C', real=True)
complsq(A*x**2 + B*x + C, x)
```

```{python error=TRUE}
complsq(x**4 + 2*x, x)
```

```{r}
eg <- function(e) {
  
  cmd <- paste0('e = ', rlang::expr_text(rlang::enexpr(e)))
  cmd <- gsub('^', '**', cmd, fixed = TRUE)
  py_run_string(cmd)
  
  m('$$\\displaystyle {{e}} \\quad=\\quad {{complsq(e, x)}}$$')
}
```

```{r}
eg(x**2 + 2*x)
```

```{r}
eg(x**2 + 2*x + 1)
```

```{r}
eg(x**2 - 2*x)
```

```{r}
eg(4*x**2 + 2*x)
```

```{r}
eg(4*x**2 - 2*x)
```

```{r}
eg(-4*x**2 + 2*x)
```

```{r}
eg(-4*x**2 - 2*x)
```

```{r}
eg(4*x**2 + 2*x + 10)
```


# Classify

## Algorithm from CEDERJ GA book

This is buggy.

```{python}
def classify_old(eq):
  
  error_msg = '''
  Equation must be of the form Ax^2 + Cy^2 + Dx + Ey + F.
  The input was ''' + latex(eq)
  
  # Used to match the coefficients of the equation
  a, b, c, d, e, f = [ \
    Wild(w, exclude=[x, y]) for w in 'abcdef' \
  ]
  
  # Match
  m = eq.match(a*x**2 + b*x*y + c*y**2 + d*x + e*y + f)
  
  # No match: bad equation
  if m is None:
    raise ValueError(error_msg)
  
  # Get coefficients
  A, B, C, D, E, F = [w.xreplace(m) for w in [a, b, c, d, e, f]]
  
  # Must be degree 2 in at least one of the vars
  # No rotated conics for now
  if B != 0 or (A == 0 and C == 0):
    raise ValueError(error_msg)
  
  # Collect everything in a dict
  retval = { \
    'eq_general': eq, \
    'A': A, \
    'B': B, \
    'C': C, \
    'D': D, \
    'E': E, \
    'F': F \
  }
  
  # Find out type of conic
  if A * C > 0:
    lmbda = D**2 / (4 * A) + E**2 / (4 * C) - F
    if A * lmbda > 0:
      if A == C:
        conic_type = 'circle'
      else:
        conic_type = 'ellipse'
    elif lmbda == 0:
      conic_type = 'point'
    else:
      conic_type = 'empty set'
  
  if A * C < 0:
    lmbda = D**2 / (4 * A) + E**2 / (4 * C) - F
    if A * lmbda != 0:
      conic_type = 'hyperbola'
    elif lmbda == 0:
      conic_type = 'concurrent lines'
  
  if A != 0 and C == 0:
    if A * E < 0:
      conic_type = 'parabola'
    elif A * E > 0:
      # Here the book had 'empty set', but the right type is 'parabola',
      # as A and E are nonzero, and the only case where this would be the 
      # empty set is if E were zero.
      conic_type = 'parabola'
    else:
      mu1 = D**2 / (4 * A) - F
      if A * mu1 >= 0:
        if 4 * F == D**2:
          conic_type = 'one line'
        else:
          conic_type = 'parallel lines'
      else:
        conic_type = 'empty set'
  
  if A == 0 and C != 0:
    if C * D < 0:
      conic_type = 'parabola'
    elif C * D > 0:
      # Here the book had 'empty set', but the right type is 'parabola',
      # as C and D are nonzero, and the only case where this would be the 
      # empty set is if D were zero.
      conic_type = 'parabola'
    else:
      mu2 = D**2 / (4 * C) - F
      if C * mu2 >= 0:
        if 4 * F == E**2:
          conic_type = 'one line'
        else:
          conic_type = 'parallel lines'
      else:
        conic_type = 'empty set'
  
  # Save type in dict
  retval['conic_type'] = conic_type
  
  return retval
```

This was being wrongly classified as the empty set. The bug was fixed:

```{python}
parab = y + x**2
parab

classify_old(parab)
```

This should be parallel lines:

```{python}
parab = y**2 - 4*y + 1
parab

classify_old(parab)
```

This should be one line:

```{python}
parab = y**2 - 4*y + 4
parab

classify_old(parab)
```

This should be the empty set:

```{python}
parab = y**2 - 4*y + 5
parab

classify_old(parab)
```

`classify()` still has a bug. Fix it.


## Analysis

### A nondegenerate parabola with $x^2$

```{python}
A, B, C, D, E, F = symbols('A B C D E F', real=True)
C = 0
A*x**2 + C*y**2 + D*x + E*y + F
```


### A degenerate parabola (only $x$ terms):

```{python}
C = E = 0

eqx = A*x**2 + D*x
a, b, c = [Wild(w, exclude=[x, y]) for w in 'abc']
eqx = complsq(eqx, x)  
a, b, c = [Wild(w, exclude=[x, y]) for w in 'abc']
m = eqx.match(a * (x + b)**2 + c)
neweq = Eq(eqx.xreplace(m), -F)
neweq
```

```{python}
a, b, c = [Wild(w) for w in 'abc']
m = neweq.lhs.match(a*b - c)
neweq = Eq(b, (c - F)/a).xreplace(m)
neweq
```

```{python}
simplify(neweq.rhs)
```

Here we can analyze the numerator of the rhs to see if any $x$ satisfies this equality:

$$
\begin{cases}
  \displaystyle
  D^2 - 4AF < 0: &\text{empty set} \\
  D^2 - 4AF = 0: &\text{one line: } & x = -D/2A \\
  D^2 - 4AF > 0: &\text{two lines: } & x = \frac{-D \pm \sqrt{D^2 - 4AF}}{2A} 
\end{cases}
$$


### A nondegenerate parabola with $y^2$

```{python}
A, B, C, D, E, F = symbols('A B C D E F', real=True)
A = 0
A*x**2 + C*y**2 + D*x + E*y + F
```


### A degenerate parabola (only $y$ terms):

```{python}
A = D = 0

eqy = C*y**2 + E*y
a, b, c = [Wild(w, exclude=[x, y]) for w in 'abc']
eqy = complsq(eqy, y)  
m = eqy.match(a * (y + b)**2 + c)
neweq = Eq(eqy.xreplace(m), -F)
neweq
```

```{python}
a, b, c = [Wild(w) for w in 'abc']
m = neweq.lhs.match(a*b - c)
neweq = Eq(b, (c - F)/a).xreplace(m)
neweq
```

```{python}
simplify(neweq.rhs)
```

Here we can analyze the numerator of the rhs to see if any $y$ satisfies this equality:

$$
\begin{cases}
  \displaystyle
  E^2 - 4CF < 0: &\text{empty set} \\
  E^2 - 4CF = 0: &\text{one line: } & y = -E/2C \\
  E^2 - 4CF > 0: &\text{two lines: } & y = \frac{-E \pm \sqrt{E^2 - 4CF}}{2C} 
\end{cases}
$$


### Combining both cases

$$
\begin{cases}
  \displaystyle
  D^2 + E^2 - 4F(A + C) < 0: &\text{empty set} 
  \\
  D^2 + E^2 - 4F(A + C) = 0: &\text{one line: } 
  & x = -D/2A &\text{or}\quad y = -E/2C 
  \\
  D^2 + E^2 - 4F(A + C) > 0: &\text{two lines: } 
  & x = \frac{-D \pm \sqrt{D^2 - 4AF}}{2A} 
  &\text{or}\quad y = \frac{-E \pm \sqrt{E^2 - 4CF}}{2C} 
\end{cases}
$$


## Algorithm 

From https://en.wikipedia.org/wiki/Degenerate_conic, considering the above analysis.

```{python}
def classify(eq):
  
  error_msg = '''
  Equation must be of the form Ax^2 + Cy^2 + Dx + Ey + F.
  The input was ''' + latex(eq)
  
  # Used to match the coefficients of the equation
  a, b, c, d, e, f = [ \
    Wild(w, exclude=[x, y]) for w in 'abcdef' \
  ]
  
  # Match
  m = eq.match(a*x**2 + b*x*y + c*y**2 + d*x + e*y + f)
  
  # No match: bad equation
  if m is None:
    raise ValueError(error_msg)
  
  # Get coefficients
  A, B, C, D, E, F = [w.xreplace(m) for w in [a, b, c, d, e, f]]
  
  # Must be degree 2 in at least one of the vars
  # No rotated conics for now
  if B != 0 or (A == 0 and C == 0):
    raise ValueError(error_msg)
  
  # Collect everything in a dict
  retval = { \
    'eq_general': eq, \
    'A': A, \
    'B': B, \
    'C': C, \
    'D': D, \
    'E': E, \
    'F': F \
  }
  
  # Discriminants
  # Non-homogeneous form
  M = Matrix([[A, B/2], [B/2, C]])
  detM = M.det()
  
  # Homogeneous form
  Q = Matrix([[A, B/2, D/2], [B/2, C, E/2], [D/2, E/2, F]])
  detQ = Q.det()
  
  # Type
  if detM > 0:
    if A == C:
      conic_type = 'circle'
    else:
      conic_type = 'ellipse'
  elif detM == 0:
    conic_type = 'parabola'
  else:
    conic_type = 'hyperbola'
    
  # Degenerate?
  is_degenerate = detQ == 0
  
  # Degenerate form
  if is_degenerate:
    
    if conic_type == 'hyperbola':
      degenerate_form = 'intersecting lines'
      
    elif conic_type == 'parabola':
      temp = D**2 + E**2 - 4 * F * (A + C)
      if temp < 0:
        degenerate_form = 'empty set'
      elif temp == 0:
        degenerate_form = 'coincident lines'
      else: # temp > 0
        degenerate_form = 'parallel lines'
        
    elif detM > 0:
      degenerate_form = 'point'
      
  else:
    degenerate_form = ''
      
  # Save info in dict
  retval['conic_type'] = conic_type
  retval['is_degenerate'] = is_degenerate
  retval['degenerate_form'] = degenerate_form
  
  return retval
```

This should be a parabola:

```{python}
parab = y + x**2
parab

classify(parab)
```

This should be parallel lines:

```{python}
parab = y**2 - 4*y + 1
parab

classify(parab)
```

This should be one line:

```{python}
parab = y**2 - 4*y + 4
parab

classify(parab)
```

This should be the empty set:

```{python}
parab = y**2 - 4*y + 5
parab

classify(parab)
```



## Tests for classify

```{python}
eq = x**2 + 2*x + y**2 + 4*y - 4
pp(classify(eq))
```

```{python error=TRUE}
eq = x + y
pp(classify(eq))
```

```{python}
eq = x**2 + x + 1
pp(classify(eq))
```

```{python error=TRUE}
eq = x * y
classify(eq)
```

```{python error=TRUE}
eq = x**2 + x*y + y**2 + x + y + 1
classify(eq)
```

```{python}
eqs = [ \
  2*x**2+y**2-4*x+5*y-1, \
  x**2+2*y**2-5*x-5*y+1, \
  -4*x**2-2*y**2-5*x+y+1, \
  -x**2-y**2-4*x+y+5, \
  -3*x**2-5*y**2+3*x-2*y, \
  5*x**2+5*y**2-5*y, \
  -x**2-2*y**2+4*x+2*y+5, \
  -2*x**2-5*y**2+y+4, \
  -x**2-y**2-4*x-3*y+4, \
  -2*x**2-4*y**2+3*x+3*y+1, \
  -5*x**2-y**2-x-4*y+2, \
  -5*x**2-2*y**2+4*x+4*y-1, \
  -4*x**2-5*y**2-2*x+2*y+2, \
  -4*x**2+y**2-4*x+5*y+4, \
  -2*x**2+2*y**2+5*x-4*y+2, \
  5*x**2-2*y**2+4*x-4*y, \
  -3*x**2+y**2-2*x+4*y+3, \
  -3*x**2+2*y**2-3*x-3*y-3, \
  -5*x**2+3*y**2+3*x+3*y-4, \
  -5*x**2+3*y**2-x+2*y-5, \
  -4*x**2+5*y**2+x-y-1, \
  5*x**2-3*y**2-5*x+4*y+2, \
  5*x**2-y**2+3*x+5*y-2, \
  -x**2+2*y**2-5*x-2*y-2, \
  4*x**2-4*y**2-x-y+1, \
  2*x**2-2*y**2+3*x+2, \
  5*y**2-3*x+3*y-2, \
  2*x**2+2*x-3*y-1, \
  4*y**2-4*x-5*y+5, \
  4*x**2-3*x-5*y-3, \
  2*y**2-4*x+5, \
  4*x**2+2*x-4*y+1, \
  -y**2+4*x+4*y-4, \
  x**2-2*x-4*y+3, \
  2*x**2-2*x-4*y+2, \
  -2*y**2+2*x-3*y-4, \
  x**2-x-4*y-2, \
  3*y**2-2*x-2*y, \
  -y**2+x+5*y+1, \
  -y**2-3*y+3, \
  -5*x**2+2*x+4, \
  2*y**2+y-1, \
  y**2+y-1, \
  x**2+4*x-1, \
  x**2+5*x+5, \
  4*y**2+y-5, \
  3*y**2-5*y-1, \
  3*x**2-5*x-5, \
  2*x**2+2*x-3, \
  x**2-4*x-4, \
  5*y**2+2*y, \
  -3*x**2-4*x, \
  -4*y**2+1, \
  5*y**2+5*y, \
  3*x**2-4*x-3, \
  4*x**2+5*x, \
  -3*y**2-y+5, \
  -x**2+5*x-3, \
  4*x**2-5*y**2+2*x+5*y-1, \
  -5*x**2+5*y**2+3*x-3*y, \
  5*x**2-5*y**2+3*x+3*y, \
  -5*x**2+5*y**2-5*x-5*y, \
  5*x**2-y**2+5*x+3*y-1, \
  4*x**2-3*y**2, \
  -4*x**2+5*y**2+2*x-5*y+1, \
  -4*x**2+4*y**2-3*x+3*y, \
  2*x**2-2*y**2-5*x-5*y, \
  x**2-y**2-2*x-2*y, \
  -x**2+4*y**2+3*x+2*y-2, \
  -3*x**2+3*y**2-4*x-2*y-1, \
  4*x**2-y**2+4*x+4*y-3, \
  x**2-y**2-2*x+1, \
  4*x**2-4*y**2-x-y, \
  -3*x**2+2*y**2+4*y+2, \
  -x**2+y**2-4*x+2*y-3, \
  2*x**2-2*y**2-3*x-5*y-2 \
]
```

```{python}
eq_list = [classify(eq) for eq in eqs]
```

```{r}
m(
  '{{ [ e[\'eq_general\'] for e in eq_list ] }} & ',
  '\\text{ {{s:[ e[\'conic_type\'] for e in eq_list ] }} } & ',
  '{{ [ e[\'is_degenerate\'] for e in eq_list ] }} & ',
  '\\text{ {{s:[ e[\'degenerate_form\'] for e in eq_list ] }} } \\\\ \n',
  before = '\\begin{array}{rlll} & \\text{Conic} & \\text{Degenerate?} & \\\\ \\hline \n',
  after = '\\hline \\end{array}'
)
```

To paste into Geogebra:

```{r}
m(
  '{{r: [ e[\'eq_general\'] for e in eq_list ]}} = 0',
  before = '{',
  after = '}',
  sep_middle = ', ',
  sep_blocks = ' ',
  use_cat = TRUE
) %>% cat()
```


# Find canonical equation

```{python}
def find_canonical_eq(conic):
  
  A, C, D, E, F = [conic[c] for c in 'ACDEF']
  conic_type = conic['conic_type']
  
  retval = None
  
  if conic_type == 'parabola':
    retval = do_parabola(A, C, D, E, F)
    
  return retval
```


## Parabola

* I have an equation

  $$
  Ax^2 + Dx + Ey + F = 0
  $$

  or
  
  $$
  Cy^2 + Dx + Ey + F = 0
  $$
  
```{python}
def do_parabola(A, C, D, E, F):
  
  # x part
  eqx = A*x**2 + D*x
  
  # y part
  eqy = C*y**2 + E*y
  
  # Used to build the new equation
  a, b, c = [Wild(w, exclude=[x, y]) for w in 'abc']
  
  if C == 0: # This is a parabola with x^2, and E != 0
    
    # Complete the square only for the x part
    eqx = complsq(eqx, x)
    # Get constants in the completed square
    m = eqx.match(a * (x + b)**2 + c)
    # Build new equation
    neweq = Eq(y + (F + c)/E, -a * (x + b)**2 / E).xreplace(m)
    
  else: # This is a parabola with y^2, and D != 0
    
    # Complete the square only for the y part
    eqy = complsq(eqy, y)
    # Get constants in the completed square
    m = eqy.match(a * (y + b)**2 + c)
    # Build new equation
    neweq = Eq(x + (F + c)/D, -a * (y + b)**2 / D).xreplace(m)
  
  return neweq
```


### Tests for parabola

```{python}
parabolas = [ \
  5*y**2-3*x+3*y-2, \
  2*x**2+2*x-3*y-1, \
  4*y**2-4*x-5*y+5, \
  4*x**2-3*x-5*y-3, \
  2*y**2-4*x+5, \
  4*x**2+2*x-4*y+1, \
  -y**2+4*x+4*y-4, \
  x**2-2*x-4*y+3, \
  2*x**2-2*x-4*y+2, \
  -2*y**2+2*x-3*y-4, \
  x**2-x-4*y-2, \
  3*y**2-2*x-2*y, \
  -y**2+x+5*y+1, \
  y + x**2, \
  x + y**2 \
]
```

```{python}
canonics = [find_canonical_eq(classify(eq)) for eq in parabolas]
```

```{r}
m(
  '{{ [ e for e in canonics ] }} \\\\',
  before = '$$\n\\displaystyle \\begin{array}{l}',
  after = '\\end{array}\n$$'
)
```

General, to paste into Geogebra:

```{r}
m(
  '{{r: [ e for e in parabolas ] }} = 0',
  sep_middle = ', ',
  before = '{',
  after = '}',
  sep_blocks = ' ',
  use_cat = TRUE  
) %>% 
  cat()
```

Canonical, to paste into Geogebra:

```{r}
m(
  '{{r: [ e.lhs for e in canonics ] }} = {{r: [ e.rhs for e in canonics ] }}',
  sep_middle = ', ',
  before = '{',
  after = '}',
  sep_blocks = ' ',
  use_cat = TRUE  
) %>% 
  cat()
```


