---
title: 'Conics'
author: 'fnaufel'
email: 'https://fnaufel.github.io/'
date: '   (v. `r format(Sys.Date(), "%d/%m/%Y")`)'
lang: 'en'

output: 
  # To install these output formats, run
  #   install.packages("devtools")
  #   devtools::install_github("fnaufel/fnaufelRmd")
  fnaufelRmd::html_report:
    []
---

```{r setup, include=FALSE}
# The next command configures MANY things and loads quite a few packages.
# 
# If you want to see what's being done, execute 
# 
#   cat(
#     system.file(
#       "rmarkdown/resources/R/_common_report.R", 
#       package = "fnaufelRmd"
#     )
#   )
# 
# to find out the location of the file. Then open the file.
# 
# If you want to change the configuration, copy the file, edit it, and
# source it instead of the package file. 
# 
# Or simply write your commands here in this code chunk.

source(
  system.file(
    "rmarkdown/resources/R/_common_report.R",
    package = "fnaufelRmd"
  )
)

library(conicas)
library(reticulate)
library(sympyglue)

sympy = import('sympy')
reticulate::py_run_file('conics.py')
```


# Complete the square

```{python}
A, B, C = symbols('A B C', real=True)
complsq(A*x**2 + B*x + C, x)
```

```{python error=TRUE}
complsq(x**4 + 2*x, x)
```

```{r}
eg <- function(e) {
  
  cmd <- paste0('e = ', rlang::expr_text(rlang::enexpr(e)))
  cmd <- gsub('^', '**', cmd, fixed = TRUE)
  py_run_string(cmd)
  
  m('$$\\displaystyle {{e}} \\quad=\\quad {{complsq(e, x)}}$$')
}
```

```{r}
eg(x**2 + 2*x)
```

```{r}
eg(x**2 + 2*x + 1)
```

```{r}
eg(x**2 - 2*x)
```

```{r}
eg(4*x**2 + 2*x)
```

```{r}
eg(4*x**2 - 2*x)
```

```{r}
eg(-4*x**2 + 2*x)
```

```{r}
eg(-4*x**2 - 2*x)
```

```{r}
eg(4*x**2 + 2*x + 10)
```


# Classify

## Analysis (parabolas)

### A nondegenerate parabola with $x^2$

```{python}
A, B, C, D, E, F = symbols('A B C D E F', real=True)
C = 0
A*x**2 + C*y**2 + D*x + E*y + F
```


### A degenerate parabola (only $x$ terms):

```{python}
C = E = 0

eqx = A*x**2 + D*x
a, b, c = [Wild(w, exclude=[x, y]) for w in 'abc']
eqx = complsq(eqx, x)  
a, b, c = [Wild(w, exclude=[x, y]) for w in 'abc']
m = eqx.match(a * (x + b)**2 + c)
neweq = Eq(eqx.xreplace(m), -F)
neweq
```

```{python}
a, b, c = [Wild(w) for w in 'abc']
m = neweq.lhs.match(a*b - c)
neweq = Eq(b, (c - F)/a).xreplace(m)
neweq
```

```{python}
simplify(neweq.rhs)
```

Here we can analyze the numerator of the rhs to see if any $x$ satisfies this equality:

$$
\begin{cases}
  \displaystyle
  D^2 - 4AF < 0: &\text{empty set} \\
  D^2 - 4AF = 0: &\text{one line: } & x = -D/2A \\
  D^2 - 4AF > 0: &\text{two lines: } & x = \frac{-D \pm \sqrt{D^2 - 4AF}}{2A} 
\end{cases}
$$


### A nondegenerate parabola with $y^2$

```{python}
A, B, C, D, E, F = symbols('A B C D E F', real=True)
A = 0
A*x**2 + C*y**2 + D*x + E*y + F
```


### A degenerate parabola (only $y$ terms):

```{python}
A = D = 0

eqy = C*y**2 + E*y
a, b, c = [Wild(w, exclude=[x, y]) for w in 'abc']
eqy = complsq(eqy, y)  
m = eqy.match(a * (y + b)**2 + c)
neweq = Eq(eqy.xreplace(m), -F)
neweq
```

```{python}
a, b, c = [Wild(w) for w in 'abc']
m = neweq.lhs.match(a*b - c)
neweq = Eq(b, (c - F)/a).xreplace(m)
neweq
```

```{python}
simplify(neweq.rhs)
```

Here we can analyze the numerator of the rhs to see if any $y$ satisfies this equality:

$$
\begin{cases}
  \displaystyle
  E^2 - 4CF < 0: &\text{empty set} \\
  E^2 - 4CF = 0: &\text{one line: } & y = -E/2C \\
  E^2 - 4CF > 0: &\text{two lines: } & y = \frac{-E \pm \sqrt{E^2 - 4CF}}{2C} 
\end{cases}
$$


### Combining both cases

$$
\begin{cases}
  \displaystyle
  D^2 + E^2 - 4F(A + C) < 0: &\text{empty set} 
  \\
  D^2 + E^2 - 4F(A + C) = 0: &\text{one line: } 
  & x = -D/2A &\text{or}\quad y = -E/2C 
  \\
  D^2 + E^2 - 4F(A + C) > 0: &\text{two lines: } 
  & x = \frac{-D \pm \sqrt{D^2 - 4AF}}{2A} 
  &\text{or}\quad y = \frac{-E \pm \sqrt{E^2 - 4CF}}{2C} 
\end{cases}
$$


## Tests

From https://en.wikipedia.org/wiki/Degenerate_conic, considering the above analysis.

```{python error=TRUE}
eq = x + y
pp(classify(eq))
```

```{python error=TRUE}
eq = x * y
classify(eq)
```

```{python error=TRUE}
eq = x**2 + x*y + y**2 + x + y + 1
classify(eq)
```

```{r echo=FALSE, eval=FALSE}
df <- conicas::gerar_conicas_df()

n <- 80
tipos <- c(
  elipse = ceiling(n/8), 
  hipérbole = ceiling(n/8),
  parábola = ceiling(n/8),
  paralelas = ceiling(n/8), 
  concorrentes = ceiling(n/8),
  ponto = ceiling(n/8),
  reta = ceiling(n/8),
  vazio = ceiling(n/8)
)

eqs <- df %>% 
  filtrar_conicas(n, tipos) %>% 
  select(tipo, eq) %>% 
  mutate(
    eq = gsub('^', '**', eq, fixed = TRUE)
  ) %>% 
  mutate(
    eq = gsub(
      '([0-9])([xy])',
      '\\1*\\2',
      eq,
      perl = TRUE
    )
  )

eqs %>% 
  pull(eq) %>% 
  substr(1, nchar(.) - 4) %>% 
  paste0(
    '  ', ., ', \\\n',
    collapse = ''
  ) %>% 
  cat()
```

```{python}
eqs = [ \
  -5*x**2-3*y**2-4*x-2*y+4, \
  4*x**2+y**2+4*x+3*y+2, \
  -5*x**2-y**2-2*x+5*y+2, \
  2*x**2+3*y**2+5*x+3, \
  x**2+5*y**2+4*x-4*y+1, \
  5*x**2+4*y**2-x-3*y-2, \
  4*x**2+2*y**2-x+3*y-3, \
  x**2+5*y**2+4*x-5*y-1, \
  -4*x**2-2*y**2-4*x+5*y+5, \
  -4*x**2-5*y**2-2*x+3*y+1, \
  2*x**2-2*y**2-4*x-4*y-2, \
  -3*x**2+y**2-4*x+5*y-3, \
  x**2-2*y**2-4*x+y+2, \
  -3*x**2+4*y**2+1, \
  -4*x**2+3*y**2+2*x+1, \
  -5*x**2+y**2+2*x-4*y+4, \
  2*x**2-5*y**2-5*x+y-2, \
  -3*x**2+y**2+4*x-4*y-1, \
  x**2-y**2-x+2*y-2, \
  3*x**2-y**2+x-3*y-5, \
  5*x**2-x-4*y+2, \
  x**2-5*x+5*y+1, \
  x**2-x-y-3, \
  3*y**2+4*x+4*y+4, \
  2*x**2+4*x-3*y-3, \
  -2*x**2-3*x-y+1, \
  2*y**2-5*x, \
  -2*x**2-2*x-4*y, \
  -5*x**2-5*x-4*y+5, \
  -3*y**2+2*x+4, \
  3*x**2+4*x+1, \
  4*y**2+2*y-4, \
  -4*x**2+x, \
  x**2-2, \
  4*y**2-y-3, \
  -3*x**2-5*x+1, \
  -x**2-2*x+3, \
  -5*y**2+3*y, \
  -4*y**2+y+5, \
  4*x**2-5*x+1, \
  -x**2+3*y**2+4*x-4, \
  5*x**2-5*y**2+4*x-4*y, \
  -4*x**2+y**2+4*x-1, \
  x**2-4*y**2-2*x+1, \
  -x**2+y**2+x+3*y+2, \
  -x**2+4*y**2+3*x+2*y-2, \
  -3*x**2+3*y**2+5*x+y-2, \
  4*x**2-5*y**2+2*x-5*y-1, \
  -x**2+5*y**2-3*x-5*y-1, \
  4*x**2-4*y**2+4*y-1, \
  -5*x**2-5*y**2-4*x+2*y-1, \
  2*x**2+y**2-4*x+2, \
  -2*x**2-4*y**2-4*y-1, \
  -x**2-4*y**2+4*x-4*y-5, \
  4*x**2+y**2+2*y+1, \
  -x**2-2*y**2+2*x-1, \
  5*x**2+y**2+4*y+4, \
  -4*x**2-2*y**2, \
  -5*x**2-5*y**2-2*x-4*y-1, \
  -3*x**2-y**2-3*x+3*y-3, \
  -5*y**2+4*y+4, \
  3*x**2, \
  -3*y**2+2*y+1, \
  -2*y**2+2*y+1, \
  -x**2+2*x+1, \
  -x**2-2*x+1, \
  -3*x**2+4*x+4, \
  -5*y**2+2*y+1, \
  x**2, \
  -5*y**2-2*y+1, \
  4*x**2+5*y**2+2*x+y+3, \
  -5*x**2-4*y**2+5*x-y-2, \
  -4*x**2-4*y**2+x+4*y-2, \
  4*x**2+4*y**2-5*x-y+5, \
  -5*x**2-3*y**2+4*x-4*y-5, \
  -5*x**2-4*y**2+2*x-y-5, \
  -2*x**2-2*y**2+2*x-5*y-4, \
  5*x**2+3*y**2-x+5*y+3, \
  5*y**2-4*y+2, \
  5*x**2+4*y**2-4*x-5*y+4 \
]
```

```{python}
eq_list = [classify(eq) for eq in eqs]
form_list = [ \
  c['degenerate_form'] if c['is_degenerate'] else '' for c in eq_list \
]
```

```{r}
m(
  '{{ [ e[\'eq_general\'] for e in eq_list ] }} & ',
  '\\text{ {{s:[ e[\'conic_type\'] for e in eq_list ] }} } & ',
  '\\text{ {{s: form_list }} } \\\\ \n',
  before = '\\begin{array}{rll} \n',
  after = '\\end{array}'
)
```

To paste into Geogebra:

Equations:

```{r}
m(
  '{{r: [ e[\'eq_general\'] for e in eq_list ]}}',
  before = '{',
  after = '}',
  sep_middle = ', ',
  sep_blocks = ' ',
  use_cat = TRUE
) %>% 
  gsub('==', '=', ., fixed = TRUE) %>% 
  cat()
```

Names:

```{r}
m(
  '"{{s: [ e[\'conic_type\'] for e in eq_list ]}}, ',
  '{{s: form_list }}"',
  before = '{',
  after = '}',
  sep_middle = ', ',
  sep_blocks = ' ',
  use_cat = TRUE
) %>% cat()
```


# Find canonical equation

## Parabola

### Nondegenerate

```{python}
parabolas = [ \
  5*y**2-3*x+3*y-2, \
  2*x**2+2*x-3*y-1, \
  4*y**2-4*x-5*y+5, \
  4*x**2-3*x-5*y-3, \
  2*y**2-4*x+5, \
  4*x**2+2*x-4*y+1, \
  -y**2+4*x+4*y-4, \
  x**2-2*x-4*y+3, \
  2*x**2-2*x-4*y+2, \
  -2*y**2+2*x-3*y-4, \
  x**2-x-4*y-2, \
  3*y**2-2*x-2*y, \
  -y**2+x+5*y+1, \
  y + x**2, \
  x + y**2 \
]
```

```{python}
parabolas = [find_canonical_eq(eq) for eq in parabolas]
pp(parabolas)
```

```{r}
m(
  '{{ [ e["eq_canonical"] for e in parabolas ] }} \\\\',
  before = '$$\n\\displaystyle \\begin{array}{l}',
  after = '\\end{array}\n$$'
)
```

General, to paste into Geogebra:

```{r}
m(
  '{{r: [ e["eq_general"] for e in parabolas ] }}',
  sep_middle = ', ',
  before = '{',
  after = '}',
  sep_blocks = ' ',
  use_cat = TRUE  
) %>% 
  gsub('==', '=', ., fixed = TRUE) %>% 
  cat()
```

Canonical, to paste into Geogebra:

```{r}
m(
  '{{r: [ e["eq_canonical"][0] for e in parabolas ] }}',
  sep_middle = ', ',
  before = '{',
  after = '}',
  sep_blocks = ' ',
  use_cat = TRUE  
) %>% 
  gsub('==', '=', ., fixed = TRUE) %>% 
  cat()
```


### Parallel lines

```{python}
parabolas = [ \
  3*x**2+4*x+1, \
  4*y**2+2*y-4, \
  -4*x**2+x, \
  x**2-2, \
  4*y**2-y-3, \
  -3*x**2-5*x+1, \
  -x**2-2*x+3, \
  -5*y**2+3*y, \
  -4*y**2+y+5, \
  4*x**2-5*x+1, \
  -5*y**2+4*y+4, \
  -3*y**2+2*y+1, \
  -2*y**2+2*y+1, \
  -x**2+2*x+1, \
  -x**2-2*x+1, \
  -3*x**2+4*x+4, \
  -5*y**2+2*y+1, \
  -5*y**2-2*y+1 \
]
```

```{python}
parabolas = [find_canonical_eq(eq) for eq in parabolas]
pp(parabolas)
```

```{r}
m(
  '{{ [ e["eq_canonical"] for e in parabolas ] }} \\\\',
  before = '$$\n\\displaystyle \\begin{array}{l}',
  after = '\\end{array}\n$$'
)
```

General, to paste into Geogebra:

```{r}
m(
  '{{r: [ e["eq_general"] for e in parabolas ] }}',
  sep_middle = ', ',
  before = '{',
  after = '}',
  sep_blocks = ' ',
  use_cat = TRUE  
) %>% 
  gsub('==', '=', ., fixed = TRUE) %>% 
  cat()
```

Canonical, to paste into Geogebra (one line):

```{r}
m(
  '{{r: [ e["eq_canonical"][0] for e in parabolas ] }}',
  sep_middle = ', ',
  before = '{',
  after = '}',
  sep_blocks = ' ',
  use_cat = TRUE  
) %>% 
  gsub('==', '=', ., fixed = TRUE) %>% 
  cat()
```

Canonical, to paste into Geogebra (other line):

```{r}
m(
  '{{r: [ e["eq_canonical"][1] for e in parabolas ] }}',
  sep_middle = ', ',
  before = '{',
  after = '}',
  sep_blocks = ' ',
  use_cat = TRUE  
) %>% 
  gsub('==', '=', ., fixed = TRUE) %>% 
  cat()
```


### Coincident lines

```{python}
parabolas = [ \
  3*x**2, \
  x**2, \
  3*y**2, \
  y**2, \
  x**2 - 4*x + 4, \
  y**2 - 4*y + 4 \
]
```

```{python}
parabolas = [find_canonical_eq(eq) for eq in parabolas]
pp(parabolas)
```

```{r}
m(
  '{{ [ e["eq_canonical"][0] for e in parabolas ] }} \\\\',
  before = '$$\n\\displaystyle \\begin{array}{l}',
  after = '\\end{array}\n$$'
)
```

General, to paste into Geogebra:

```{r}
m(
  '{{r: [ e["eq_general"] for e in parabolas ] }}',
  sep_middle = ', ',
  before = '{',
  after = '}',
  sep_blocks = ' ',
  use_cat = TRUE  
) %>% 
  gsub('==', '=', ., fixed = TRUE) %>% 
  cat()
```

Canonical, to paste into Geogebra:

```{r}
m(
  '{{r: [ e["eq_canonical"] for e in parabolas ] }}',
  sep_middle = ', ',
  before = '{',
  after = '}',
  sep_blocks = ' ',
  use_cat = TRUE  
) %>% 
  gsub('==', '=', ., fixed = TRUE) %>% 
  cat()
```


### Empty set

```{python}
parabolas = [ \
  5*y**2 - 4*y + 2, \
  5*x**2 - 4*x + 2, \
]
```

```{python}
parabolas = [find_canonical_eq(eq) for eq in parabolas]
pp(parabolas)
```


## Ellipse or circle

### Analysis

```{python}
A, C, D, E, F = symbols('A C D E F', real=True)
eq = A*x**2 + C*y**2 + D*x + E*y + F

eqx = complsq(A*x**2 + D*x, x)
eqy = complsq(C*y**2 + E*y, y)

eqsq = eqx + eqy + F

eqsq
```

```{python}
a2 = Wild('a2', exclude=[x, y])
b2 = Wild('b2', exclude=[x, y])
sq1 = Wild('sq1', exclude=[y])
sq2 = Wild('sq2', exclude=[x])
ind = Wild('ind', exclude=[x, y])

m = eqsq.match(a2 * sq1 + b2 * sq2 - ind)

eqcan = sq1 / (ind / a2) + sq2 / (ind / b2)
eqcan = eqcan.subs(m)
eqcan
```

This sum must equal $1$.

When is this impossible?

We know (because this is an ellipse) that $A$ and $C$ have the same sign.

So, either both terms in the sum above are positive or both are negative.

This is the empty set iff both are negative, i.e., iff

$$
A \cdot \left(`r m('{{ ind.subs(m) }}')`\right) < 0
$$

I must test for this in the ellipse case.


### Tests

```{python}
eqs = [ \
  -5*x**2-5*y**2-4*x-2*y+4, \
  4*x**2+4*y**2+4*x+3*y+2, \
  -5*x**2-5*y**2-2*x+5*y+2, \
  2*x**2+3*y**2+5*x+3, \
  x**2+5*y**2+4*x-4*y+1, \
  5*x**2+4*y**2-x-3*y-2, \
  4*x**2+2*y**2-x+3*y-3, \
  x**2+5*y**2+4*x-5*y-1, \
  -4*x**2-2*y**2-4*x+5*y+5, \
  -4*x**2-5*y**2-2*x+3*y+1, \
  -5*x**2-5*y**2-4*x+2*y-1, \
  2*x**2+y**2-4*x+2, \
  -2*x**2-4*y**2-4*y-1, \
  -x**2-4*y**2+4*x-4*y-5, \
  4*x**2+y**2+2*y+1, \
  -x**2-2*y**2+2*x-1, \
  5*x**2+y**2+4*y+4, \
  -4*x**2-2*y**2, \
  -5*x**2-5*y**2-2*x-4*y-1, \
  -3*x**2-y**2-3*x+3*y-3, \
  4*x**2+5*y**2+2*x+y+3, \
  -5*x**2-4*y**2+5*x-y-2, \
  -4*x**2-4*y**2+x+4*y-2, \
  4*x**2+4*y**2-5*x-y+5, \
  -5*x**2-3*y**2+4*x-4*y-5, \
  -5*x**2-4*y**2+2*x-y-5, \
  -2*x**2-2*y**2+2*x-5*y-4, \
  5*x**2+3*y**2-x+5*y+3, \
  5*x**2+4*y**2-4*x-5*y+4 \
]
```

```{python}
ellipses = [find_canonical_eq(c) for c in eqs]
form_list = [ \
  c['degenerate_form'] if c['is_degenerate'] else '' for c in ellipses \
]
```

```{r}
py$ellipses %>% 
  map_chr(~pluck(., 'conic_type'))
```

General, to paste into Geogebra:

```{r}
m(
  '{{r: [ e["eq_general"] for e in ellipses ] }}',
  sep_middle = ', ',
  before = '{',
  after = '}',
  sep_blocks = ' ',
  use_cat = TRUE  
) %>% 
  gsub('==', '=', ., fixed = TRUE) %>% 
  cat()
```

Canonical, to paste into Geogebra:

```{r}
m(
  '{{r: [ e["eq_canonical"][0] for e in ellipses ] }}',
  sep_middle = ', ',
  before = '{',
  after = '}',
  sep_blocks = ' ',
  use_cat = TRUE  
) %>% 
  gsub('==', '=', ., fixed = TRUE) %>% 
  cat()
```

```{r}
m(
  '{{ [ e[\'eq_general\'] for e in ellipses ] }} & ',
  '\\displaystyle {{ [ e[\'eq_canonical\'] for e in ellipses ] }} & ',
  '\\text{ {{s: form_list }} }\\\\ \n',
  before = '\\begin{array}{rll} \n',
  after = ' \\end{array}'
)
```


## Hyperbola

### Tests

```{python}
eqs = [ \
  2*x**2-2*y**2-4*x-4*y-2, \
  -3*x**2+y**2-4*x+5*y-3, \
  x**2-2*y**2-4*x+y+2, \
  -3*x**2+4*y**2+1, \
  -4*x**2+3*y**2+2*x+1, \
  -5*x**2+y**2+2*x-4*y+4, \
  2*x**2-5*y**2-5*x+y-2, \
  -3*x**2+y**2+4*x-4*y-1, \
  x**2-y**2-x+2*y-2, \
  3*x**2-y**2+x-3*y-5, \
  -x**2+3*y**2+4*x-4, \
  5*x**2-5*y**2+4*x-4*y, \
  -4*x**2+y**2+4*x-1, \
  x**2-4*y**2-2*x+1, \
  -x**2+y**2+x+3*y+2, \
  -x**2+4*y**2+3*x+2*y-2, \
  -3*x**2+3*y**2+5*x+y-2, \
  4*x**2-5*y**2+2*x-5*y-1, \
  -x**2+5*y**2-3*x-5*y-1, \
  4*x**2-4*y**2+4*y-1, \
]
```

```{python}
hyps = [find_canonical_eq(c) for c in eqs]
form_list = [ \
  c['degenerate_form'] if c['is_degenerate'] else '' for c in hyps \
]
```

General, to paste into Geogebra:

```{r}
m(
  '{{r: [ e["eq_general"] for e in hyps ] }}',
  sep_middle = ', ',
  before = '{',
  after = '}',
  sep_blocks = ' ',
  use_cat = TRUE  
) %>% 
  gsub('==', '=', ., fixed = TRUE) %>% 
  cat()
```

Canonical, to paste into Geogebra:

First eq:

```{r}
m(
  '{{r: [ e["eq_canonical"][0] for e in hyps ] }}',
  sep_middle = ', ',
  before = '{',
  after = '}',
  sep_blocks = ' ',
  use_cat = TRUE  
) %>% 
  gsub('==', '=', ., fixed = TRUE) %>% 
  cat()
```

```{r}
m(
  '{{ [ e[\'eq_general\'] for e in hyps ] }} & ',
  '\\displaystyle {{ [ e[\'eq_canonical\'] for e in hyps ] }} & ',
  '\\text{ {{s: form_list }} }\\\\ \n',
  before = '\\begin{array}{rll} \n',
  after = ' \\end{array}'
)
```


# Find center, foci, etc.

## Circles

```{python}
def analyze_circle(conic):
  
  eq = conic['eq_canonical'][0]
  eqlhs = eq.lhs
  eqrhs = eq.rhs
  
  h, k, r = [Wild(w, exclude=[x, y]) for w in 'hkr']
  
  mlhs = eqlhs.match((x - h)**2 + (y - k)**2)
  mrhs = eqrhs.match(r**2)
  
  conic['center'] = Point(h.xreplace(mlhs), k.xreplace(mlhs))
  conic['radius'] = r.xreplace(mrhs)
```

```{python}
eq = -5*x**2-5*y**2-4*x-2*y+16
conic = find_canonical_eq(eq)
analyze_circle(conic)
pp(conic)
```

## Ellipses

```{python}
def analyze_ellipse(conic):
  
  eq = conic['eq_canonical'][0]
  # I can ignore the 1 on the rhs
  eqlhs = eq.lhs
  
  h, k, a, b = [Wild(w, exclude=[x, y]) for w in 'hkab']
  
  # Will args always have the x term first and the y term second?
  # Careful to avoid evaluation
  eqx, eqy = sympify('eqlhs.args', locals={'eqlhs': eqlhs}, evaluate=False)
  mx = eqx.match((1 / a**2) * (x - h)**2)
  my = eqy.match((1 / b**2) * (y - k)**2)
  
  aa = a.xreplace(mx)
  bb = b.xreplace(my)
  
  major_axis = 'h'
  
  if (aa < bb):
    aa, bb = bb, aa
    major_axis = 'v'
  
  hh = h.xreplace(mx)
  kk = k.xreplace(my)
  
  conic['center'] = Point(hh, kk)
  conic['a'] = aa
  conic['b'] = bb
  conic['c'] = sqrt(aa**2 - bb**2)
  conic['eccentricity'] = cc / aa
  
  if major_axis == 'h':
    conic['focus'] = [ \
      Point(hh - cc, kk), Point(hh + cc, kk) \
    ]
    ...
    
  
```

```{python}
eq = 2*x**2+3*y**2+5*x+3
conic = find_canonical_eq(eq)
analyze_ellipse(conic)
pp(conic)
```
